---
title: "decathlon"
output: html_document
---

```{r}
library(tidyverse)
```


#read in the raw data about decathlon
```{r}
library(here)

decathlon_data <- read_rds(here("raw_data/decathlon.rds"))

#look at the overview
glimpse(decathlon_data)
```

# convert row names to a column with values
```{r}
library(tibble)
decathlon_data_clean <- rownames_to_column(decathlon_data, var = "names")
```

# clean column names
```{r}
library(janitor)

decathlon_data_clean <- decathlon_data_clean %>% 
  clean_names() %>% 
  rename("100m" = "x100m", "400m" = "x400m", "110m_hurdle" = "x110m_hurdle", "1500m" = "x1500m")


decathlon_data_clean
```
# convert all names to lowercase
```{r}
decathlon_data_clean <- decathlon_data_clean %>% 
 mutate(names = tolower(names))
```

# pivot table to longer format
```{r}
decathlon_data_long <- decathlon_data_clean %>% 
  pivot_longer(2:11, names_to = "event", values_to = "result")

decathlon_data_long
```
# check for NAs in the data
```{r}
decathlon_data_long %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```

# check for zeros in the result column
```{r}
decathlon_data_long %>% 
  filter(result == 0)
```

# Who had the longest long jump seen in the data?
```{r}
#filter by event and get the highest number in result
longest_jump <- decathlon_data_long %>% 
  filter(event == "long_jump") %>% 
  slice_max(result, n = 1)

longest_jump 
```

# What was the average 100m time in each competition?
```{r}
#group by competition and filter by event to get the average for both competition
avg_100m <- decathlon_data_long %>% 
  group_by(competition) %>% 
  filter(event == "100m") %>% 
  summarise(mean(result)) 

avg_100m
```

# Who had the highest total points across both competitions?
```{r}
# group by name
# get total points for everyone 
# pull out highest total_points
highest_total_points <- decathlon_data_long %>% 
  group_by(names) %>% 
  summarise(total_points = sum(points)) %>% 
  select(names, total_points) %>% 
  ungroup() %>% 
  slice_max(total_points, n =1)

highest_total_points
```

# What was the shot-put scores for the top three competitors in each competition?

```{r}
#top 3 competitors in each competition

#first get top 3 competitors in Decaster and create a vector
decaster_top3 <- decathlon_data_long %>% 
  group_by(names) %>% 
  filter(competition == "Decastar") %>% 
  filter(rank == c("1", "2", "3")) %>% 
  select("names") %>% 
  distinct() %>% 
  pull()

#first get top 3 competitors in OlympicG and create a vector
olympicg_top3 <- decathlon_data_long %>% 
  group_by(names) %>% 
  filter(competition == "OlympicG") %>% 
  filter(rank == c("1", "2", "3")) %>% 
  select("names") %>% 
  distinct() %>% 
  pull()

#filter byy results in above code  and then filter on shot put scores
shot_put_scores_top_3 <- decathlon_data_long %>% 
  filter(names %in% decaster_top3) %>% 
  filter(event == "shot_put")

shot_put_scores_top_3
```

#What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r}
# get the average points for competitors who ran the 400m in less than 50 seconds
avg_points_400m_under_50s <- decathlon_data_long %>% 
  filter(event == "400m", 
         result < "50") %>% 
  summarise(mean(points)) %>% 
  pull()
  
# get the average points for competitors who ran the 400m in more than 50 seconds
avg_points_400m_over_50s <- decathlon_data_long %>% 
  filter(event == "400m", 
         result > "50") %>% 
  summarise(mean(points)) %>% 
  pull()

# calculate the difference
difference_in_avg_400m <- 
  sum(avg_points_400m_under_50s - avg_points_400m_over_50s)

difference_in_avg_400m
```

