---
title: "halloween_candy_analysis"
author: "Barry O Dea"
date: "14/07/2021"
output: html_document
---

# This file is for performing analysis on the Halloween Candy Data in it's clean and tidy version. 

# 1. Install packages required 
```{r}
library(tidyverse)
```


# 2. Read in the clean csv file
```{r}
library(here)
candy_rating_top_countries <- read_csv(here("clean_data/candy_rating_top_countries.csv"))
```

# 3. Perform Analysis
# Q1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)?
```{r}
# group data by year to get the ratings per year and not overall and filter out where there is no ratings
rating_count <- candy_rating_top_countries %>% 
  group_by(year) %>% 
  count(rating) %>% 
  filter(rating != "no_rating") %>% 
  ungroup()

# summarise the number of ratings per year
ratings_per_year <- rating_count %>% 
  group_by(year) %>% 
  summarise(sum(n))
```



# Q2. What was the average age of people who are going out trick or treating?
```{r}
# assume the ages over 100 are an error and filter out and any age of 0 is ommitted as these are mainly where no age was given
# first filter by those going out
mean_age_going_out <- candy_rating_top_countries %>% 
  filter(age > 0, age < 100) %>% 
  filter(going_out == "Yes") %>% 
  summarise(mean(age)) %>% 
  pull()
```

# Q3. What was the average age of people who are not going trick or treating?
```{r}
# then my those who are not going out
mean_age_not_going_out <- candy_rating_top_countries %>% 
  filter(age > 0, age < 100) %>% 
  filter(going_out == "No") %>% 
  summarise(mean(age)) %>% 
  pull()
```



# For each of joy, despair and meh, which candy bar revived the most of these ratings?
```{r}
top_rating_per_candy <- candy_rating_top_countries %>% 
  group_by(rating) %>% 
  count(candy) %>% 
  slice_max(n, n = 1)

top_rating_per_candy
```

# Q4. How many people rated Starburst as despair?
```{r}
starburst_despair_count <- candy_rating_top_countries %>% 
  filter(candy == "starburst",
         rating == "DESPAIR") %>% 
  count(rating) 

starburst_despair_count
```

# For the next three questions, count despair as -1, joy as +1 and meh as 0.
```{r}
# create a new column converting rating values into numeric values
candy_rating_nums <- candy_rating_top_countries %>% 
  filter(rating != "no_rating") %>% 
  mutate(rating_num = 
           case_when(
             rating == "DESPAIR" ~ -1,
             rating == "JOY" ~ 1,
             rating == "MEH" ~ 0
           ))
```


# Q.5 What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_rating_by_gender <- candy_rating_nums %>% 
  group_by(gender, candy) %>% 
  summarise(rating_num = sum(rating_num)) %>% 
  slice_max(rating_num, n = 1)

candy_rating_by_gender
```


# Q6. What was the most popular candy bar in each year?
```{r}
candy_rating_by_year <- candy_rating_nums %>% 
  group_by(year, candy) %>% 
  summarise(rating_num = sum(rating_num)) %>% 
  slice_max(rating_num, n = 1)

candy_rating_by_year 
```


# Q7. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?

```{r}
candy_rating_by_country<- candy_rating_nums %>% 
  group_by(country, candy) %>% 
  summarise(rating_num = sum(rating_num)) %>% 
  slice_max(rating_num, n = 1)

candy_rating_by_country 
```



